---
output: github_document
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Edit README.Rmd, then render it. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
devtools::load_all(".", quiet = TRUE)
```

# fhktools

[![R-CMD-check](https://github.com/JINGYIYIYIYI/fhktools/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/JINGYIYIYIYI/fhktools/actions/workflows/R-CMD-check.yaml)
[![pkgdown](https://github.com/JINGYIYIYIYI/fhktools/actions/workflows/pkgdown.yaml/badge.svg)](https://github.com/JINGYIYIYIYI/fhktools/actions/workflows/pkgdown.yaml)
[![MIT license](https://img.shields.io/badge/license-MIT-blue.svg)](LICENSE.md)

`fhktools` is a defensive, dependency-free R implementation of the standard
Foster--Haltiwanger--Krizan (FHK) productivity decomposition. It is designed
for empirical panels where duplicated units, malformed shares, industry-code
changes, survey rotation, missing values, and ambiguous entry/exit definitions
are real concerns.

中文用户请参阅[中文完整教程](vignettes/fhk-zh.Rmd)。An English step-by-step
guide is available in the [introductory vignette](vignettes/fhk-introduction.Rmd).

## Why this package?

Many short FHK scripts reproduce the five-term accounting formula but silently
make consequential data decisions. `fhktools` separates three questions:

1. **Does the accounting identity close?** Every result reports a scaled
   residual and `identity_ok`.
2. **Are the input shares and panel keys valid?** Unsafe factor-to-numeric
   conversions, duplicated unit-period rows, negative weights, missing group
   labels, and malformed shares are rejected by default.
3. **What does turnover mean?** Without registry information, entry and exit
   are explicitly labeled as sample presence. Firm birth/death, sample
   turnover, and movement between industries are kept conceptually separate.

The endpoint identity is

\[
\Delta P = \text{Within} + \text{Between} + \text{Cross}
           + \text{Entry} + \text{Exit},
\]

with

\[
\begin{aligned}
\text{Within}  &= \sum_{i\in C}s_{i0}(p_{i1}-p_{i0}),\\
\text{Between} &= \sum_{i\in C}(s_{i1}-s_{i0})(p_{i0}-P_0),\\
\text{Cross}   &= \sum_{i\in C}(s_{i1}-s_{i0})(p_{i1}-p_{i0}),\\
\text{Entry}   &= \sum_{i\in E}s_{i1}(p_{i1}-P_0),\\
\text{Exit}    &= -\sum_{i\in X}s_{i0}(p_{i0}-P_0).
\end{aligned}
\]

The returned `exit` component already contains the leading minus sign.

## Installation

Install the development version from GitHub:

```{r installation, eval=FALSE}
install.packages("pak")
pak::pak("JINGYIYIYIYI/fhktools")

# Or:
install.packages("remotes")
remotes::install_github("JINGYIYIYIYI/fhktools")
```

## Quick start with raw weights

The package includes a synthetic, unbalanced firm panel with genuine endpoint
entry and exit, stable industry codes, raw employment, global shares,
within-industry shares, and registry-style activity windows.

```{r data-preview}
library(fhktools)
data(fhk_example)
head(fhk_example[, c("firm_id", "year", "industry", "productivity", "employment")])
```

Raw sales, employment, or output weights are usually safest because the package
can construct period-specific shares with a transparent denominator.

```{r quick-start}
overall <- fhk_decomposition(
  fhk_example,
  firm_col = "firm_id",
  time_col = "year",
  prod_col = "productivity",
  weight_col = "employment",
  t0 = 2010,
  t1 = 2012
)

round(overall[, c(
  "aggregate_change", "within", "between", "cross", "entry", "exit",
  "residual"
)], 6)
```

`aggregate_change` equals the sum of the five components up to floating-point
rounding. Percentage columns are signed contributions relative to the signed
aggregate change. They are `NA` when the aggregate change is too close to zero.

## Validate genuine firm entry and exit

Sample absence can reflect business closure, survey rotation, matching failure,
or a missing observation. If registry-style activity dates are available, use
them explicitly:

```{r activity-window}
validated <- fhk_decomposition(
  fhk_example,
  "firm_id", "year", "productivity", "employment",
  2010, 2012,
  first_active_col = "first_active_year",
  last_active_col = "last_active_year",
  turnover_validation = "error"
)

validated[, c(
  "n_continuers", "n_entrants", "n_exiters",
  "turnover_basis", "turnover_validated"
)]
```

Without those columns, `turnover_basis` is `"sample_presence"`. That is an
interpretation warning, not an accounting failure.

## Grouped decompositions

```{r grouped}
by_industry <- fhk_decomposition(
  fhk_example,
  "firm_id", "year", "productivity", "employment",
  2010, 2012,
  group_cols = "industry"
)

by_industry[, c(
  "industry", "aggregate_change", "within", "between", "cross",
  "entry", "exit", "identity_ok"
)]
```

Within-group weights are normalized inside each group. Do **not** add the
industry components to construct an economy-wide decomposition. Use
`fhk_hierarchical()` to obtain both levels from correctly defined weights:

```{r hierarchical}
both_levels <- fhk_hierarchical(
  fhk_example,
  "firm_id", "year", "productivity", "employment",
  2010, 2012,
  group_cols = "industry"
)

both_levels$overall[, c("aggregate_change", "identity_ok")]
both_levels$by_group[, c("industry", "aggregate_change", "identity_ok")]
```

## Pre-computed shares

The package distinguishes shares that sum to one globally from shares that sum
to one inside every group-period:

```{r share-scopes}
from_global_shares <- fhk_decomposition(
  fhk_example,
  "firm_id", "year", "productivity", "global_share",
  2010, 2012,
  group_cols = "industry",
  weight_is_share = TRUE,
  share_scope = "global"
)

from_within_group_shares <- fhk_decomposition(
  fhk_example,
  "firm_id", "year", "productivity", "within_industry_share",
  2010, 2012,
  group_cols = "industry",
  weight_is_share = TRUE,
  share_scope = "within_group"
)
```

Malformed shares fail by default. Set `renormalize_input_shares = TRUE` only
when renormalization is substantively justified; the result records which
periods were renormalized.

## Complex panels: explicit policies

`fhk_complex_example` deliberately contains an industry mover, a group present
at only one endpoint, a zero-weight row, and a missing alternative productivity
value.

```{r complex-preview}
data(fhk_complex_example)
subset(
  fhk_complex_example,
  firm_id %in% c("B", "J") | employment == 0 | is.na(productivity_alt)
)[, c("firm_id", "year", "industry", "productivity_alt", "employment")]
```

The default grouped call stops because a continuing firm changes industry. If
the movement is real and the old/new industry interpretation is intended, make
that policy explicit. Endpoint-only groups can likewise be dropped explicitly;
they remain available in an audit attribute.

```{r complex-policy}
complex_result <- fhk_decomposition(
  fhk_complex_example,
  "firm_id", "year", "productivity", "employment",
  2010, 2012,
  group_cols = "industry",
  first_active_col = "first_active_year",
  last_active_col = "last_active_year",
  group_change_action = "allow",
  incomplete_group_action = "drop"
)

attr(complex_result, "skipped_groups")
complex_result[, c(
  "industry", "n_group_movers", "turnover_basis",
  "global_firm_turnover_validated", "identity_ok"
)]
```

An industry mover is an old-industry exit and a new-industry entrant, but it is
not a firm death or birth. The `global_firm_turnover_*` columns preserve the
separate firm-level activity audit.

## Adjacent periods

```{r adjacent}
annual <- fhk_adjacent(
  fhk_example,
  "firm_id", "year", "productivity", "employment",
  expected_step = 1
)

annual[, c("period_start", "period_end", "aggregate_change", "identity_ok")]
```

`expected_step = 1` prevents 2010--2012 from being silently treated as an
annual transition when 2011 is absent.

## Multiple productivity measures on a common sample

```{r multiple}
multiple <- fhk_multi_measure(
  fhk_example,
  "firm_id", "year",
  c("productivity", "productivity_alt"),
  "employment",
  2010, 2012,
  common_sample = TRUE
)

multiple[, c("measure", "aggregate_change", "identity_ok")]
```

With `common_sample = TRUE` and `na_action = "drop"`, a row missing any requested
measure is removed once for all measures. This avoids comparing decompositions
that use different firm samples.

## What the package checks

| Risk | Default behavior | Explicit alternative |
|---|---|---|
| Factor productivity or weights | Error | Convert and verify outside the function |
| Duplicate unit-period rows | Error | Aggregate or construct a valid unit ID |
| Missing/non-finite required values | Error | `na_action = "drop"` with warning |
| Negative weights | Error | None |
| Zero weights | Keep and report | `zero_weight_action = "drop"` or `"error"` |
| Shares do not sum to one | Error | `renormalize_input_shares = TRUE` |
| Continuing unit changes group | Error | `group_change_action = "warn"` or `"allow"` |
| Group absent at one endpoint | Error | `incomplete_group_action = "warn"` or `"drop"` |
| Reversed endpoints | Error | `require_t1_after_t0 = FALSE` |
| Accounting identity does not close | Error | Inspect with `fail_on_identity_error = FALSE` |

## Interpretation limits

Software cannot choose the empirical estimand for you. Before reporting a
decomposition, decide and document:

- whether the analytical unit is a firm, establishment, or firm-product;
- whether employment, sales, or output is the economically appropriate weight;
- whether productivity is in levels or logs;
- whether absence means true exit, sample rotation, or failed matching;
- whether industry classifications are harmonized across time; and
- whether a long difference or annual transitions match the research question.

A perfectly closing identity does not prove that these choices are correct.

## Citation

```{r citation}
citation("fhktools")
```

The methodological reference is:

Foster, L., Haltiwanger, J., and Krizan, C. J. (2001). “Aggregate
Productivity Growth: Lessons from Microeconomic Evidence.” In *New
Developments in Productivity Analysis*. <https://doi.org/10.3386/w6803>.

## Development

```{r development, eval=FALSE}
devtools::test()
devtools::check()
pkgdown::build_site()
```

Contributions are welcome. Please read [CONTRIBUTING.md](CONTRIBUTING.md) and do
not upload confidential firm-level data.
