---
title: "A Practical Guide to FHK Productivity Decomposition"
author: "Yu Li"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A Practical Guide to FHK Productivity Decomposition}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(fhktools)
```

## 1. What is being decomposed?

For a unit-level productivity measure $p_{it}$ and period-specific weight share
$s_{it}$, aggregate productivity is

$$P_t = \sum_i s_{it}p_{it}.$$

Between endpoints 0 and 1, units are classified as continuers ($C$), entrants
($E$), or exiters ($X$). `fhktools` implements

$$
\begin{aligned}
P_1-P_0 =
&\sum_{i\in C}s_{i0}(p_{i1}-p_{i0}) \\
&+\sum_{i\in C}(s_{i1}-s_{i0})(p_{i0}-P_0) \\
&+\sum_{i\in C}(s_{i1}-s_{i0})(p_{i1}-p_{i0}) \\
&+\sum_{i\in E}s_{i1}(p_{i1}-P_0) \\
&-\sum_{i\in X}s_{i0}(p_{i0}-P_0).
\end{aligned}
$$

These lines correspond to within, between, cross, entry, and exit. The returned
`exit` column already contains the minus sign.

The identity is algebraic. A small residual verifies implementation and input
consistency, but does not prove that productivity, weights, unit identifiers, or
turnover are economically well defined.

## 2. The minimum data contract

Each endpoint panel should have:

1. one row per analytical unit and period;
2. a stable identifier across time;
3. a numeric productivity measure;
4. a non-negative raw size or share; and
5. an ordered, non-factor period variable.

```{r clean-data}
data(fhk_example)
str(fhk_example)
```

The analytical unit matters. If a firm can own several plants, a firm ID cannot
be used with plant-level productivity unless the plants are first aggregated or
a plant ID is used. The package checks duplicated unit-period rows before any
group split, including duplicates that occur under different industry labels.

## 3. Start with raw weights

Using raw employment, sales, or output is generally safer than creating shares
outside the function. The denominator then remains explicit in the result.

```{r raw-weights}
z <- fhk_decomposition(
  fhk_example,
  firm_col = "firm_id",
  time_col = "year",
  prod_col = "productivity",
  weight_col = "employment",
  t0 = 2010,
  t1 = 2012
)

z[, c(
  "P0", "P1", "aggregate_change",
  "within", "between", "cross", "entry", "exit",
  "component_sum", "residual", "identity_ok"
)]
```

Important audit columns include:

- `weight_total_t0` and `weight_total_t1`;
- counts and weight shares of continuers, entrants, and exiters;
- `turnover_basis` and `turnover_validated`; and
- `residual` and `identity_ok`.

## 4. Raw weights versus pre-computed shares

The clean example includes global employment shares. They reproduce the raw
weight calculation:

```{r raw-share-equivalence}
from_share <- fhk_decomposition(
  fhk_example,
  "firm_id", "year", "productivity", "global_share",
  2010, 2012,
  weight_is_share = TRUE
)

fields <- c("aggregate_change", "within", "between", "cross", "entry", "exit")
rbind(raw = unlist(z[fields]), supplied_share = unlist(from_share[fields]))
```

When `weight_is_share = TRUE`, shares must sum to one at the declared scope.
Automatic renormalization is off by default because it can conceal a wrong
denominator. If renormalization is substantively justified, set
`renormalize_input_shares = TRUE` and retain the audit fields in the archived
results.

## 5. Sample turnover versus genuine birth and death

A two-endpoint panel identifies presence and absence. It does not automatically
identify business birth and death. Survey rotation, intermittent reporting, and
failed linkage can all mimic entry or exit.

The example supplies registry-style activity windows:

```{r validated-turnover}
validated <- fhk_decomposition(
  fhk_example,
  "firm_id", "year", "productivity", "employment",
  2010, 2012,
  first_active_col = "first_active_year",
  last_active_col = "last_active_year",
  turnover_validation = "error"
)

validated[, c(
  "n_continuers", "n_entrants", "n_exiters",
  "share_entrants_t1", "share_exiters_t0",
  "turnover_basis", "turnover_validated"
)]
```

Entrants must have first activity in $(t_0,t_1]$; exiters must have last activity
in $[t_0,t_1)$. Continuers must be active by $t_0$ and through $t_1$.

If no registry fields exist, report the terms as *sample entry* and *sample
exit*. Do not relabel them as business creation and destruction.

## 6. Grouped analysis

```{r grouped-analysis}
g <- fhk_decomposition(
  fhk_example,
  "firm_id", "year", "productivity", "employment",
  2010, 2012,
  group_cols = "industry"
)

g[, c(
  "industry", "P0", "P1", "aggregate_change",
  "within", "between", "cross", "entry", "exit"
)]
```

Weights are normalized within every group-period. These components describe
within-industry productivity dynamics; they are not additive across industries.
Use `fhk_hierarchical()` when both levels are needed:

```{r hierarchical-analysis}
h <- fhk_hierarchical(
  fhk_example,
  "firm_id", "year", "productivity", "employment",
  2010, 2012,
  group_cols = "industry"
)

h$overall[, c("aggregate_change", "identity_ok")]
h$by_group[, c("industry", "aggregate_change", "identity_ok")]
```

If only within-industry shares are available, industry weights in the aggregate
economy cannot be recovered. Raw weights or global shares are needed for a
hierarchical analysis.

## 7. Industry movers and endpoint-only groups

`fhk_complex_example` is designed to trigger safeguards:

```{r complex-data}
data(fhk_complex_example)
subset(
  fhk_complex_example,
  firm_id %in% c("B", "J") | employment == 0 | is.na(productivity_alt)
)[, c("firm_id", "year", "industry", "productivity_alt", "employment")]
```

Firm B changes industry, and Construction appears only at the first endpoint.
The default grouped analysis stops. If the industry movement is genuine and the
old-group exit/new-group entry interpretation is intended, say so explicitly:

```{r explicit-complex-policy}
g_complex <- fhk_decomposition(
  fhk_complex_example,
  "firm_id", "year", "productivity", "employment",
  2010, 2012,
  group_cols = "industry",
  first_active_col = "first_active_year",
  last_active_col = "last_active_year",
  group_change_action = "allow",
  incomplete_group_action = "drop"
)

attr(g_complex, "skipped_groups")
g_complex[, c(
  "industry", "n_group_movers", "turnover_basis",
  "global_firm_turnover_validated", "identity_ok"
)]
```

The group-specific `turnover_validated` remains false when movers are present,
because firm activity windows cannot validate entry into or exit from a group.
The separate `global_firm_turnover_validated` field records the firm birth/death
audit.

When industry-code changes are administrative rather than economic, harmonize
codes before decomposition instead of allowing movers.

## 8. Missing values and zero weights

Missing or non-finite required values stop by default. `na_action = "drop"`
issues a warning because dropping a row can create an apparent entrant or
exiter. Archive the warning and compare sample counts before and after any
drop-based robustness specification.

Zero-weight rows are retained and counted by default. They do not affect the
weighted aggregate but do affect presence-based classifications. Alternatives
are `zero_weight_action = "drop"` and `zero_weight_action = "error"`. A period
or group with a zero total weight cannot be decomposed.

## 9. Annual and other adjacent transitions

```{r adjacent-transitions}
annual <- fhk_adjacent(
  fhk_example,
  "firm_id", "year", "productivity", "employment",
  expected_step = 1
)

annual[, c(
  "period_start", "period_end", "aggregate_change",
  "n_entrants", "n_exiters", "identity_ok"
)]
```

The wrapper uses global observed periods, so groups share common endpoints.
`expected_step` protects against treating the next observed year as the next
calendar year when there is a gap.

Long differences and annual decompositions answer different questions. A firm
that enters after 2010 and exits before 2012 is absent from both long-difference
endpoints but can contribute to annual transitions.

## 10. Multiple productivity measures

```{r multiple-measures}
m <- fhk_multi_measure(
  fhk_example,
  "firm_id", "year",
  c("productivity", "productivity_alt"),
  "employment",
  2010, 2012,
  common_sample = TRUE
)

m[, c("measure", "aggregate_change", "within", "entry", "exit")]
```

For measures with different missingness, use `common_sample = TRUE` and
`na_action = "drop"` to remove rows invalid for any requested measure once.
Otherwise, apparently different decompositions may reflect different samples.

## 11. Reporting checklist

A reproducible empirical table should report or archive:

1. analytical unit and identifier construction;
2. endpoint periods and whether the comparison is annual or a long difference;
3. productivity level/log definition and estimation method;
4. raw weighting variable and denominator scope;
5. group classification and crosswalk policy;
6. sample-entry/exit versus registry-validated birth/death interpretation;
7. missing-value, zero-weight, mover, and incomplete-group policies;
8. continuer, entrant, and exiter counts and shares; and
9. the accounting residual and `identity_ok`.

Save the complete returned data frame before selecting columns for publication.
The audit fields are part of the empirical record.

## 12. Citation

```{r citation}
citation("fhktools")
```

The methodological reference is Foster, Haltiwanger, and Krizan (2001),
“Aggregate Productivity Growth: Lessons from Microeconomic Evidence,”
<https://doi.org/10.3386/w6803>.
